---
title: "LDT_final"
author: "Hadil"
date: "2023-02-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
install.packages("optimx")
install.packages("lme4")
install.packages("Matrix")
install.packages("ggplot2")
install.packages("usethis")
install.packages("tidyverse")
install.packages("nlme")
install.packages("janitor")
devtools::install_github('cttobin/ggthemr')
devtools::install_github("hadley/devtools")

```


```{r}
library(Matrix)
library(readxl)
library(usethis)
library(ggplot2)
library(ggthemr)
library(lme4)
library(devtools)
library(Matrix)
library(DT)
library(optimx)
library(ggdist)
library(brms)
library(ggeffects)
library(dplyr)
library(tidyverse)
library(nlme)
library(lmerTest)
```


## coding level factors
```{r}
LDT_final$Condition = ifelse(LDT_final$condition4 == "CONG", -0.5, 0.5)
```

```{r}
LDT_final <- LDT_final %>%
  mutate(hs1 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "HS1" ~ 1L,  # target
                          TRUE      ~ 0L), # anything else
         hs2 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "HS2" ~  1L, # target
                          TRUE      ~ 0L), # anything else
          rt1 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "RT1" ~  1L, # target
                          TRUE      ~ 0L), # anything else
          rt2 = case_when(group == "AHS" ~ -1L, # baseline
                          group == "RT2" ~  1L, # target
                          TRUE      ~ 0L))

```

```{r}
LDT_final <- LDT_final %>%
  mutate(incong = case_when(condition == "Congruent" ~ -1L, # baseline
                          condition == "Incongruent" ~  1L, # target
                          TRUE      ~ 0L), # anything else
          unacw = case_when(condition == "Congruent" ~ -1L, # baseline
                          condition == "UnaccWithArEq" ~  1L, # target
                          TRUE      ~ 0L), # anything else
          unacwt = case_when(condition == "Congruent" ~ -1L, # baseline
                          condition == "UnaccWithoutArEq" ~  1L, # target
                          TRUE      ~ 0L))

```

```{r}
library(dplyr)

sum_table_LDT_accuracy <- LDT_final %>%
  mutate(accuracy = as.numeric(as.character(accuracy))) %>%
  group_by(condition, group) %>%
  summarise(
    Total = mean(accuracy)
  ) %>%
  ungroup()

print(sum_table_LDT_accuracy)
```
```{r}
library(dplyr)

sum_table_LDT_rt <- LDT_final %>%
  group_by(condition, group) %>%
  summarise(
    Total = mean(rt, na.rm = TRUE)
  ) %>%
  ungroup()

print(sum_table_LDT_rt)
```


```{r}
LDT_final$group <- as.factor(LDT_final$group)
LDT_final$item <- as.factor(LDT_final$item)
LDT_final$accuracy <- as.factor(LDT_final$accuracy)

str(LDT_final)
```
```{r}
LDT_final$Dominance_score <- scale(LDT_final$Dominance_score, center = TRUE, scale = TRUE)
LDT_final$English_use <- scale(LDT_final$English_use, center = TRUE, scale = TRUE)
LDT_final$Arabic_use <- scale(LDT_final$Arabic_use, center = TRUE, scale = TRUE)
colnames(LDT_final)[30] <- "Domscale"
colnames(LDT_final)[22] <- "Englishscale"
colnames(LDT_final)[23] <- "Arabicscale"
LDT_final$English_attidutes <- scale(LDT_final$English_attidutes, center = TRUE, scale = TRUE)
LDT_final$Arabic_attidutes <- scale(LDT_final$Arabic_attidutes, center = TRUE, scale = TRUE)
LDT_final$Arabic_history <- scale(LDT_final$Arabic_history, center = TRUE, scale = TRUE)
LDT_final$English_history <- scale(LDT_final$English_history, center = TRUE, scale = TRUE)
colnames(LDT_final)[26] <- "Englishattscale"
colnames(LDT_final)[27] <- "Arabicattscale"
colnames(LDT_final)[20] <- "Enghistscale"
colnames(LDT_final)[21] <- "Arabhistscale"



```


```{r}
ggthemr("fresh")
library(readxl)
LDT_final <- read_excel("LDT_final.xlsm")
View(LDT_final)
  janitor::clean_names() |>
  mutate(participant = as.factor(participant),
         group = as.factor(group),
  )

knitr::opts_chunk$set(echo = TRUE)
```
```{r}
LDT_final %>%
  group_by(group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),     # Add standard deviation calculation
         range = ~diff(range(.x, na.rm = TRUE)))   # Add range calculation
  ))
```
```{r}
LDT_final %>%
  group_by(group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),     # Add standard deviation calculation
         range = ~diff(range(.x, na.rm = TRUE)),   # Add range calculation
         min = ~min(.x, na.rm = TRUE),     # Add minimum calculation
         max = ~max(.x, na.rm = TRUE))     # Add maximum calculation
  ))
```

```{r}
LDT_final %>%
  group_by(group) %>%
  group_by(condition, group) %>%
  summarise(across(
    where(is.numeric),
    list(mean = ~mean(.x, na.rm = TRUE),
         count = ~sum(!is.na(.x)),
         sd = ~sd(.x, na.rm = TRUE),
         range = ~diff(range(.x, na.rm = TRUE)),
         min = ~min(.x, na.rm = TRUE),     # Add minimum calculation
         max = ~max(.x, na.rm = TRUE)      # Add maximum calculation
    ))) 
```

```{r}
hist(LDT_final$rt)
log_rt_LDT <- log(LDT_final$rt)
hist(log_rt_LDT)
```

# RT per condition

```{r}
LDT_final |>
  ggplot(aes(x = group, y = log_rt_LDT, fill = condition)) +
  geom_boxplot()
```
```{r}
LDT_final |>
  mutate(accuracy = as.numeric(as.character(accuracy))) %>%
  ggplot(aes(x = condition, y = accuracy, fill = group)) +
  geom_boxplot()
```
## Data summary

```{r}
LDT_final %>%
group_by(group) %>%
  summarise(across(where(is.numeric),
    list(mean = ~ mean(.x, na.rm = TRUE), count = ~ sum(!is.na(.x)))))
```
```{r}
library(dplyr)

sum_table_acc <- LDT_final %>%
  mutate(accuracy = as.numeric(as.character(accuracy))) %>%
  group_by(group) %>%
  summarise(
    Total = mean(accuracy)
  ) %>%
  ungroup()

print(sum_table_acc)
```

# Correct answers per group & condition

```{r}
library(data.table)
freqs <- LDT_final |>
  group_by(group, condition) |>
  summarize(total_correct = sum(accuracy),
            fract_correct = total_correct / n()) |>
  arrange(condition, group)

freqs |> data.table()


```
# Correct answers per group & condition, visualized as bar plot

```{r, fig.width=20, fig.height=6}
freqs |>
  ggplot(aes(x = group, y =accuracy, fill = group)) +
  geom_col() +
  facet_wrap(~ condition, ncol = 4) +
  theme(legend.position = "bottom")
```

# RTs per group & condition

```{r}
rts <- LDT_final|>
  filter(!is.na(log_rt_LDT)) |>
  group_by(group, condition) |> 
  summarize(median_rt = median(log_rt_LDT),
            iqr_rt = IQR(log_rt_LDT)) |> 
  arrange(condition, group)

rts |> data.table()
```


# RTs visualized using boxplots

```{r, fig.width=20, fig.height=6}

LDT_final |>
  filter(!is.na(log_rt_LDT)) |> 
  ggplot(aes(x = group, y = log_rt_LDT, fill = group)) +
  geom_boxplot() +
  facet_wrap(~ condition, ncol = 4) +
  theme(legend.position = "bottom")

```

# Anomalous participant: unusually large number of observations

```{r}
LDT_final|> 
  group_by(participant, condition) |> 
  count() |> 
  pull(n) |> 
  map_lgl(\(x) x == 23) |> 
  all()
```

# Models (Accuracy) (Binomial Logistic Regression)

```{r}
# mod0 <- glmer(as.factor(key_resp_5.CORR) ~ group + (1|group/participant) + (1|condition), family = binomial, data = GapTaskLong, control = glmerControl(check.conv.grad = .makeCC("warning", tol = 5e-2)))

lbfgs_ctrl <- glmerControl()

xls_nb <- LDT_final |>
  mutate(condition = as.factor(condition) |> relevel(ref = "Congruent"),
         group = as.factor(group) |> relevel(ref = "AHS"),
         correct = accuracy == 1)

mod_0_LDT <- glmer(accuracy ~ 1 + (1|participant),
               family = binomial(link = "logit"), 
               data = LDT_final,
               control = lbfgs_ctrl)

mod_1_LDT <- glmer(accuracy ~ 1 + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final,
               control = lbfgs_ctrl)

mod_2_LDT <- glmer(accuracy ~ group + (1|participant),
               family = binomial(link = "logit"),
               data = LDT_final,
               control = lbfgs_ctrl)

mod_3_LDT <- glmer(accuracy ~ group + (1|participant) + (1|item),
               family = binomial(link = "logit"),
               data = LDT_final,
               control = lbfgs_ctrl)

mod_4_LDT <- glmer(accuracy ~ group * condition + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)


print(anova(mod_0_LDT, mod_1_LDT, mod_2_LDT, mod_3_LDT,mod_4_LDT ))
```

```{r}
print(summary(mod_3_LDT))
```
##Model estimates (Accuracy with group)
This clearly indicates a significant main effect is shown between the returnees and the AHS.


```{r}
model_params_LDT <- broom.mixed::tidy(mod_3_LDT)

model_params_LDT|>
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  labs(y = "Logit", x = "Group")
```
# Here I attempt to perform post-hoc by using emmeans for mod_3 (accuracy)

```{r}
library(emmeans)
emmeans(mod_3_LDT, list(pairwise ~ group), adjust = "tukey")
```

# Models with RT

```{r}
mod0_rt_LDT <- lmer(log_rt_LDT ~ 1 + (1|participant), data = LDT_final)

mod1_rt_LDT <- lmer(log_rt_LDT ~ 1 + (1|participant) + (1|item), data = LDT_final)

mod2_rt_LDT <- lmer(log_rt_LDT ~ group + (1|participant), data = LDT_final)

mod3_rt_LDT <- lmer(log_rt_LDT ~ group + (1|participant) + (1|item), data = LDT_final)

mod4_rt_LDT <- lmer(log_rt_LDT ~ group + length + (1|participant) + (1|item), data = LDT_final)


print(anova(mod0_rt_LDT, mod1_rt_LDT, mod2_rt_LDT, mod3_rt_LDT, mod4_rt_LDT))
```

```{r}
summary(mod4_rt_LDT)
```


```{r}

ranef(mod3_rt)
hist(ranef(mod3_rt)$participant[,1])
hist(ranef(mod3_rt)$item[,1])
hist(resid(mod3_rt))
```

# In the graph below it clearly shows a significant main effect between the AHS and the other groups.

```{r}
model_params_rt_LDT <- broom.mixed::tidy(mod3_rt_LDT)

model_params_rt_LDT |> 
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  labs(y = "Estimate", x = "Group")
```

# Here I attempt to perform post-hoc by using emmeans for  (reaction time residuals)

```{r}
library(emmeans)
emmeans(mod4_rt_LDT, list(pairwise ~ group), adjust = "tukey")
```
#Models with interaction with condition for ACCURACY

```{r}
settingAGQ=0
```

##interaction between group and condition for ACCURACY
##condition and group as a fixed effects and a random intercept for participants and for items. “congruent” condition was the reference level for condition.
```{r}



mod_0cd_LDT <- glmer(accuracy ~ 1 + (1|participant), 
               family = binomial(link = "logit"), 
               data = LDT_final,
             nAGQ=settingAGQ,control = lbfgs_ctrl) 

mod_1cd_LDT <- glmer(accuracy ~ 1 + (1|participant) + (1|item), 
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control= lbfgs_ctrl)

mod_2cd_LDT <- glmer(accuracy ~ group + (1|participant), 
               family = binomial(link = "logit"), 
               data = LDT_final,
              nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_3cd_LDT <- glmer(accuracy ~ group + (1|participant) + (1|item), 
               family = binomial(link = "logit"), 
               data = LDT_final,
             nAGQ=settingAGQ,control  = lbfgs_ctrl)

mod_4cd_LDT<- glmer(accuracy ~ group * condition + (1|participant),
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_5cd_LDT <- glmer(accuracy ~ group * condition + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)


mod_6cd_LDT <- glmer(accuracy ~ group * condition + length + group*length + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_7cd_LDT <- glmer(accuracy ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_8cd_LDT <- glmer(accuracy ~ (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final,
               nAGQ=settingAGQ,control = lbfgs_ctrl)



print(anova(mod_0cd_LDT, mod_1cd_LDT, mod_2cd_LDT, mod_3cd_LDT, mod_4cd_LDT, mod_5cd_LDT, mod_6cd_LDT, mod_7cd_LDT, mod_8cd_LDT))
```

mod_9cd_LDT <- glmer(accuracy ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1+condition|participant) + (1|item),
               family = binomial(link = "logit"), 
               data = LDT_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_10cd_LDT <- glmer(accuracy ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (hs1 + hs2 + rt1 + rt2) * Arabicscale + (hs1 + hs2 + rt1 + rt2) * Englishscale + (1|participant) + (1|item), family = binomial(link = "logit"), data = LDT_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

mod_11cd_LDT <- glmer(accuracy ~ group + (hs1 + hs2 + rt1 + rt2) * condition + (hs1 + hs2 + rt1 + rt2) * Arabicscale + (hs1 + hs2 + rt1 + rt2) * Englishscale + (1|participant) + (1|item), family = binomial(link = "logit"), data = LDT_final, nAGQ=settingAGQ,control = lbfgs_ctrl)

```{r}
print(summary(mod_8cd_LDT))
```

```{r}
model_params <- broom.mixed::tidy(mod_9cd_LDT)
model_params |>
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate, color = term)) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.3, color = "black") +
  geom_point(size = 2) +
  scale_color_manual(values = c("red", "blue", "green", "orange", "purple", "pink", "yellow", "maroon", "black", "brown", "beige", "grey", "magenta", "cyan","seagreen" "skyblue", "darkgray", "navyblue", "lightgrey","darkmagenta" )) +  # Adjust the colors as needed
  theme(legend.position = "bottom") +
  labs(y = "Accuracy", x = "Group")
```


```{r}
library(emmeans)
emmeans(mod_5cd_LDT, list(pairwise ~ group), adjust = "tukey")
```
# Model estimates for Accuracy

```{r}
model_params_cd_LDT <- broom.mixed::tidy(mod_5cd_LDT)

model_params_cd_LDT |> 
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  labs(y = "Logit", x = "Group") +
  scale_color_viridis_d() +
  coord_flip()

```

# RT (interaction with condition)

```{r}

mod_0_rtcd_LDT <- lmer(log_rt_LDT ~ 1 + (1|participant), 
               data = LDT_final)

mod_1_rtcd_LDT <- lmer(log_rt_LDT ~ group + (1|participant), 
               data = LDT_final)

mod_2_rtcd_LDT<- lmer(log_rt_LDT ~ group + (1|participant) + (1|item), data = LDT_final)

mod_3_rtcd_LDT <- lmer(log_rt_LDT ~ group * condition + (1|participant) + (1|item), data = LDT_final)

mod_4_rtcd_LDT <- lmer(log_rt_LDT ~ group * length + (1|participant) + (1|item), data = LDT_final)

mod_5_rtcd_LDT <- lmer(log_rt_LDT ~ length + group + (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item), data = LDT_final)

mod_6_rtcd_LDT <- lmer(log_rt_LDT ~ length + (hs1 + hs2 + rt1 + rt2) * condition + (1|participant) + (1|item), data = LDT_final)


mod_7_rtcd_LDT <- lmer(log_rt_LDT ~ length + group + (hs1 + hs2 + rt1 + rt2) * condition + (hs1 + hs2 + rt1 + rt2) * Arabicscale + (hs1 + hs2 + rt1 + rt2) * Englishscale + (1|participant) + (1|item), data = LDT_final)

mod_8_rtcd_LDT <- lmer(log_rt_LDT ~ length + group * condition + (1|participant) + (1|item) + (1+condition|participant), data = LDT_final)



anova(mod_0_rtcd_LDT, mod_1_rtcd_LDT, mod_2_rtcd_LDT, mod_3_rtcd_LDT, mod_4_rtcd_LDT, mod_5_rtcd_LDT, mod_6_rtcd_LDT, mod_7_rtcd_LDT, mod_8_rtcd_LDT)
```


##According to the results, it shows that the t values are all below two so there is no main effect of RT (interaction with condition)


```{r}
summary(mod_8_rtcd_LDT)
```
```{r}
library(ggplot2)
ggplot(LDT_final, aes(x = accuracy, y = rt, color = group)) +
  geom_point() +
  geom_smooth(method='lm',formula=y~x)
```



```{r}
anova(mod_6_rtcd)
library(lmerTest)
```

```{r}
model_params_rtcd <- broom.mixed::tidy(mod_5_rtcd)

model_params_rtcd |> 
  filter(!str_detect(term, "sd")) |>
  ggplot(aes(x = term, y = estimate)) +
  # geom_hline(yintercept = 1, linetype = "dotted", color = "black") +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.1, color = "black") +
  geom_point(aes(color = term), size = 3) +
  theme(legend.position = "bottom") +
  scale_color_viridis_d() +
  labs(y = "Estimate", x = "Group") +
  coord_flip()
```
